---
title: "The Kalman filter and the square root Kalman filter"
author: "Koichi (Koiti) Yano"
format: pdf
editor: visual
---

## Introduction

## The Kalman filter and the square root Kalman filter

### A linear Gaussian state space model

A linear Gaussian state space model is defined by the following two equations.

$$
x_t=Fx_{t-1} + E u_t + w_t
$$

and

$$y_t=Hx_t+v_t,$$

where $x_t$ is the $(k \times 1)$ state vector , $y_t$ is the $(l \times 1)$ observation vector, $u_t$ is the $(n \times 1)$ exogenous vector, $v_t$ is the $(k \times 1)$ state noise, $w(t)$ is the $(l \times 1)$ observation noise, $t$ is a time index. The Kalman Filter is used to estimate the state vector $x_t$ given the observations $y_t$.

The Kalman filter is implemented using the following recursion:

1.  Prediction step:

    1.  $x_{t|t-1} = F x_{t-1|t-1} + E u_t$

    2.  $P_{t|t-1} = F P_{t-1|t-1} F^t + V$

2.  Innovation step:

    1.  $e_t = y_t - H x_{t|t-1}$

    2.  $s_t = H P_{t-1|t-1}H^t + W$

3.  Update step:

    1.  $K_t=P_{t|t-1} H^t s_t^{-1}$

    2.  $x_{t|t} = x_{t|t-1} + K_t e_t$

    3.  $P_{t|t} = (I - K_t H) P_{t|t-1}$

where $x_{t|t}$ is the estimated state vector at, $P_t$ is the estimated state covariance matrix, and $K_t$ is the Kalman gain at time t. See Kitagawa, (2010) for more details.

### The square root Kalman filter using only QR decompositions

The square root Kalman filter is an alternative to the Kalman filter that is numerically more stable. The square root Kalman filter is implemented using the following recursion. Tracy (2022) proposes the following recursion for the square root Kalman filter using only QR decompositions (The QR Kalman filter).

The QR Kalman filter is implemented using the following recursion

1.  Prediction step

    1.  $x_{t|t-1} = F x_{t-1|t-1} + E u_t$

    2.  $\Sigma_{t|t-1} = gr_r(\Sigma_{t-1|t-1} F^t, \Gamma_v)$

2.  Innovation step:

    1.  $e_t = y_t - H x_{t|t-1}$

    2.  $G_t = gr_r (\Sigma_{t|t-1} H^t, \Gamma_w)$

3.  Update step:

    1.  $K_t={[G_t^{-1} (G_t^{-t} H) {\Sigma_{t|t-1}}^t(\Sigma_{t|t-1})]}^t$

    2.  $x_{t|t} = x_{t|t-1} + K_t e_t$

    3.  $\Sigma_{t|t} = gr_r(\Sigma(t|t-1) {(I - K(t) H)}^t, \Gamma_w {K(t)}^t)$

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
